---
title: "demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = whitebox::check_whitebox_binary(),
  collapse = TRUE,
  comment = "#>",
  progress = FALSE
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

<!-- $$(K_{macro} -  K_{FieldCond}) {\frac{W_{surf}-Field_{cap}}{d_{soil}}} ({\frac{Sat_{mc} - Field_{cap}}{d_{soil} + K_{FieldCond}}})^{-1}$$ -->
<!-- $${\frac{W_{surf}-Field_{cap}}{d_{soil}}}$$ -->
<!-- $$ {\frac{Sat_{mc} - Field_{cap}}{d_{soil} + K_{FieldCond}}$$ -->
The purpose of this package is to provide a method of computational hydrology for arid regions.
Arid regions provide unique challenge to hydrological modeling ranging from 
very steep terrain with canyons and cliffs, lack of detailed soil data and land cover data,
complex interactions between infiltration and lack of detailed rainfall data.

The utility of the desertHydro model is built upon the integration of field measurements, e.i.,
rainfall and stream gauge data, with remotely sensed or publically available data.

Some of the data that can be utilized within this package:
Rainfall data: https://www.goes-r.gov/
Some information about GOES files: https://docs.opendata.aws/noaa-goes16/cics-readme.html#about-the-data


## Load in the necessary libraries
```{r, setup, message=FALSE, warning=FALSE}
library(desertHydro)
library(terra)
library(ggplot2) # For plotting
library(gganimate) # For animations
library(whitebox)
#require(whitebox)
whitebox::wbt_init(
  exe_path = whitebox::wbt_exe_path(shell_quote = FALSE),
  check_version = TRUE
)

```

## Model Set-up
The desertHydro package is split into three levels.

### 1 Pre-process: Create and gather necessary files
Model Folder -> where to save the outputs from the model
WatershedElements Folder -> folder containing the necessary inputs for the model run.


```{r, important folders, results='hide'}
# Important steps

#ModelFolder <- "ModelFolder" # replace path to folder where model is saved
ModelFolder <- tempdir()
#ModelFolder <- r"(C:\PackageDev\desertHydro\ModelElementsTest)" # <- Example

#WatershedElements <- "./inst/extdata/DemoElements" # replace this with your watershed
WatershedElements <- dirname(system.file("extdata", "dem.tif", package = "desertHydro")) # replace this with your watershedelements folder
# Checking model folder
# WatershedElements <- r"(C:\PackageDev/desertHydro/inst/extdata/DemoElements/)" # replace this with your watershed elements folder

```

The priority of the model is to utilize the intersection of land cover types, 
soil profiles from soil databases, and adjustable tables (excel files) to model
a particular rainfall event. 

Necessary Inputs:
1) DEM or digital elevation file
2) Manning's n or surface roughness map (.tif or .shp) embedded within land cover file
With these two elements, the model can run on impervious mode (impervious = T)

```{r, important files, results='hide'}

demFile <- "dem.tif" # DEM file to simulate over - should be named smooth/dem.tif

# Check if watershed elements folder exists
WatershedElements <- folderCheck(WatershedElements, demFile)

boundary <- "test-boundary.shp" # Name of shapefile for outline
landCoverFile <- "soils.shp" # name of soils data
LandCoverCharacteristics <- "LandCoverCharacteristics_Soils.xlsx"
key <- "MUSYM" # key is present in both the landcoverfile and the land cover characteristics
```

## Note: 
One of the key aspects of the map is to ensure that the soil characteristics
can be cross referenced to an input data table.
The three things that need to be present when adding soil characteristics.
### 1) Land cover spatial file: .shp or .tif
### 2) Land cover characteristics: excel file containing soil depth characteristics
### 3) Key: a character string that is present in both the attribute table of the spatial file
and the excel file.

### Rainfall event elements
```{r, event files, results='hide'}
date <- "2018-07-13" # Optional - if dates present must be in YYYY-MM-DD format
rainFile <- "USGS_Rain_2001_2021.xlsx"
dischargeFile <- "observable-discharge.csv"
discharge = T # If TRUE, expects discharge present in file

# If rainfall method is synthetic - uses 1/2 inch - 15 minute storm
rainfall_method <- "synthetic" 
```

### Model elements
```{r, model checklist, results='hide'}

length <- 10 # grid cell length - assumes uniform grid
store = T
time_step <- .1 # time step in minutes
simulation_length <-  1 # time length in minutes
impervious <- T  # no infiltration
overwrite <- T # overwrite watershed elements and model folder
write <- T # create graphs and write outputs to model folder
restartModel <- F # If model is paused mid-run, can be restarted with same inputs
gif <- F # If TRUE, gif's will be created automatically. Can be created at the end

```

### Check if necessary files are present
```{r, check files}
# Check if the necessary folders are present
foldersToCheck <- c(WatershedElements)
folders <- sapply(foldersToCheck, FUN = file.exists)
if(!all(folders)){
  print(paste0("All folders not found. List of folders not found."))
  print(folders)
}
# Check if the necessary files are present
filesToCheck <- c(demFile, boundary, landCoverFile, 
                  LandCoverCharacteristics, rainFile, dischargeFile)
filesCheck <- sapply(file.path(WatershedElements, filesToCheck), 
                     FUN = file.exists)
if(!all(filesCheck)){
  print("All files are not found. Here is a list of the files present.")
  print(filesCheck)
}else{
  print("All files found.")
}
```

## 2 Model Run - off and running
```{r, model run, echo = T, results='hide', cache = T}

a <- arid_model(ModelFolder,
                WatershedElements,
                date = date,
                demFile = demFile,
                boundary = boundary,
                landCoverFile = landCoverFile,
                LandCoverCharacteristics = LandCoverCharacteristics,
                key = key,
                impervious = impervious,
                rainfall_method = rainfall_method,
                store = store,
                gif = gif,
                discharge = discharge,
                time_step = time_step,
                simulation_length = simulation_length,
                overwrite = overwrite,
                write = write,
                restartModel = restartModel
                )
SoilStack <- terra::rast(file.path(ModelFolder, "model_soil_stack.tif"))
```

```{r}

# Plot the maximum values for each cell
surfaceStorage <- terra::rast(file.path(ModelFolder, "surfaceStorage.tif"))
maxDepth <- max(surfaceStorage)
terra::writeRaster(maxDepth, file.path(ModelFolder, "max-depth-map.tif"))

#plot(maxDepth)
velocityStorage <- terra::rast(file.path(ModelFolder, "velocityStorage.tif"))
maxVelocity <- max(velocityStorage)
terra::writeRaster(maxVelocity, file.path(ModelFolder, "max-velocity-map.tif"))
#plot(maxVelocity)
slope <- SoilStack$slope
dem <- terra::rast(file.path(ModelFolder, "model_dem.tif"))
# Plot values that are a little to high - maybe
plot(maxDepth > 30)
plot(maxVelocity > 4)
terra::writeRaster(maxDepth > 30, file.path(ModelFolder, "greaterThan30cm.tif"))
terra::writeRaster(maxVelocity > 4, file.path(ModelFolder, "greaterThan4ms.tif"))

# stream section
stream <- terra::vect(file.path(WatershedElements, "stream_path.shp"))
# Load in created mainstem in QGIS
# ms <- terra::vect(file.path(WatershedElements, "main-stem.shp"))
# crsAssign(file.path(WatershedElements, "main-stem.shp"))
stream <- 
# dem <- terra::rast(breach_test)
model_dem_rast <- terra::rast(file.path(ModelFolder, "model_dem.tif"))
ogdem <- terra::rast(file.path(WatershedElements, "dem.tif"))

plot(ogdem - model_dem_rast)
dem <- model_dem_rast
# Extract values along stream
demExt <- terra::extract(dem, stream)
depthExt <- terra::extract(maxDepth, stream, xy = T)
velExt <- terra::extract(maxVelocity, stream)
slopeExt <- terra::extract(slope, stream)
stream_x <- nrow(demExt)
x <- 1:stream_x
vValues <- c(75, 97, 112, 138, 147, 194, 200)
# Combine datasets
df <- data.frame(x = x, dem = demExt$model_dem)

demG <- ggplot(df) + geom_line(aes(x, dem)) + geom_vline(xintercept = vValues, color = "grey")

df <- data.frame(x = x, depth = depthExt$max, velocity = velExt$max, slope = slopeExt$slope, dem = demExt$model_dem)
# Plot values
labelx <- "Distance from Gauge (# cells)"

velocityG <- ggplot(df) + geom_point(aes(x, velocity), colour = "red", shape = "square") +
  labs(title = "Canyon Stream Properties") + xlab("") + ylab("Velocity (m/s)")+ theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept = vValues, color = "grey")

depthG <- ggplot(df) + geom_point(aes(x, depth), color = "blue", shape = "circle")+xlab("") + ylab("Maximum depth (cm)") + theme_minimal() + geom_vline(xintercept = vValues, color = "grey")

slopeG <- ggplot(df) + geom_point(aes(x, slope), color = "black", shape = "diamond")+xlab(labelx) + ylab("Slope (°)") + theme_minimal() + geom_vline(xintercept = vValues, color = "grey")

demG <- ggplot(df) + geom_line(aes(x, dem))+ geom_vline(xintercept = vValues, color = "grey")
library(gridExtra)

comb <- gridExtra::grid.arrange(velocityG, depthG, slopeG, demG)
ggsave(file.path(ModelFolder, "stream_props.png"), plot = comb, width = 8, height = 8)

plot(x, depthExt$max)

```


```{r gif-creation, gganimate = list(nframes = 30, fps = 3)}
# Rain file to grab the necessary rainfall method
rain_file <- desertHydro:::rainfallMethodCheck(ModelFolder, rainfall_method)
gifs <- desertHydro:::gifCreation(ModelFolder, rain_file = rain_file,
                                  rainfall_method = rainfall_method, discharge = discharge,
                                  saveGraph = T, date = date)
```
```{r, discharge}
# Determine discharge peak and volume for a given day
dischargeDF <- dailyDischarge(discharge_file_path = file.path(WatershedElements, "example_discharge.csv"),
                                    discharge_date = date,
                                    save_location = ModelFolder,
                                    saveGraphs = F)
volume <- dischargeVolume(dischargeDF)

```
```{r}
# Flow to string
flow_accum <- terra::rast(file.path(WatershedElements, "flow_accumulation.tif"))

# Extract mainstep - 300,000 cells
main_stem <- terra::ifel(flow_accum > 200000, 1, 0)
writeRaster(main_stem, file.path(WatershedElements, "main_stem.tif"))
main_lines <- as.lines(main_stem)


plot(main_stem)

# Vectorize mainstem
```
