---
title: "watershedWorkflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{watershedWorkflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  progress = FALSE
  
)
```

<!-- $$(K_{macro} -  K_{FieldCond}) {\frac{W_{surf}-Field_{cap}}{d_{soil}}} ({\frac{Sat_{mc} - Field_{cap}}{d_{soil} + K_{FieldCond}}})^{-1}$$ -->
<!-- $${\frac{W_{surf}-Field_{cap}}{d_{soil}}}$$ -->
<!-- $$ {\frac{Sat_{mc} - Field_{cap}}{d_{soil} + K_{FieldCond}}$$ -->
## Load in the necessary libraries
```{r, setup, message=FALSE, warning=FALSE}
library(desertHydro)
library(terra)
library(ggplot2) # For plotting
library(gganimate) # For animations
require(whitebox)
whitebox::wbt_init() # initiation whitebox

```

## Model Set-up
The desertHydro package is split into three levels.

### 1 Pre-process: Create and gather necessary files
Model Folder -> where to save the outputs from the model
WatershedElements Folder -> folder containing the necessary inputs for the model run.


```{r, important folders, results='hide'}
# Important steps

ModelFolder <- "../ModelElementsTest" # replace path to folder where model is saved
#ModelFolder <- r"(C:\PackageDev\desertHydro\ModelElementsTest)"

WatershedElements <- "../inst/extdata/DemoElements" # replace this with your watershed elements folder
# WatershedElements <- r"(C:\PackageDev/desertHydro/inst/extdata/DemoElements/)" # replace this with your watershed elements folder

```

The priority of the model is to utilize the intersection of land cover types, 
soil profiles from soil databases, and adjustable tables (excel files) to model
a particular rainfall event. 

Necessary Inputs:
1) DEM or digital elevation file
2) Manning's n or surface roughness map (.tif or .shp) embedded within land cover file
With these two elements, the model can run on impervious mode (impervious = T)

```{r, important files, results='hide'}

demFile <- "dem.tif" # DEM file to simulate over
boundary <- "boundary.shp" # Optional
landCoverFile <- "soils.shp"
LandCoverCharacteristics <- "LandCoverCharacteristics_soils.xlsx"
key <- "MUSYM" # key is present in both the landcoverfile and the land cover characteristics
```

## Note: One of the key aspects of the map is to ensure that the soil charactertistics
can be cross referenced to an input data table.
The three things that need to be present when adding soil characteristics.
1) Land cover spatial file: .shp or .tif
2) Land cover characteristics: excel file containing soil depth characteristics
3) Key: a character string that is present in both the attribute table of the spatial file
and the excel file.

### Rainfall event elements
```{r, event files, results='hide'}
date <- NULL # Optional - if dates present must be in YYYY-MM-DD format
rainFile <- "USGS_Rain_2001_2021.xlsx"
dischargeFile <- "example_discharge.csv"
discharge = F # If TRUE, expects discharge present in file

# If rainfall method is synthetic - uses 1/2 inch - 15 minute storm
rainfall_method <- "Synthetic" 
```

### Model elements
```{r, model checklist, results='hide'}

length <- 10 # grid cell length - assumes uniform grid
store = T
time_step <- .1 # time step in minutes
simulation_length <-  15 # time length in minutes
impervious <-  # no infiltration
overwrite <- T # overwrite previous inputs in both watershed elements and model folder
write <- T # create graphs and write outputs to model folder
restartModel <- F # If model is paused mid-run, can be restarted with same inputs
gif <- F # If TRUE, gif's will be created automatically. Can be created at the end
```

### Check if necessary files are present
```{r, check files}
# Check if the necessary folders are present
foldersToCheck <- c(ModelFolder, WatershedElements)
folders <- sapply(foldersToCheck, FUN = file.exists)
if(!all(folders)){
  print(paste0("All folders are not found. Here is a list of folders not found."))
  print(folders)
}
# Check if the necessary files are present
filesToCheck <- c(demFile, boundary, landCoverFile, LandCoverCharacteristics, rainFile, dischargeFile)
filesCheck <- sapply(file.path(WatershedElements, filesToCheck), FUN = file.exists)
if(!all(filesCheck)){
  print("All files are not found. Here is a list of the files present.")
  print(filesCheck)
}else{
  print("All files found.")
}
```

## 2 Model Run - off and running
```{r, model run, echo = F, results='hide', cache = T}

a <- arid_model(ModelFolder,
                WatershedElements,
                date = date,
                demFile = demFile,
                boundary = boundary,
                landCoverFile = landCoverFile,
                LandCoverCharacteristics = LandCoverCharacteristics,
                key = key,
                impervious = impervious,
                rainfall_method = rainfall_method,
                store = store,
                gif = gif,
                discharge = discharge,
                time_step = time_step,
                simulation_length = simulation_length,
                overwrite = overwrite,
                write = write,
                restartModel = restartModel
                )

```
The model will create files within the Model Folder was it runs. The length of simulation
is based on the size of the grid, the initial timestep, and the duration. It is always encouraged
to run a small scale version in both space and time to determine if things are working properly.

## 3 Post processing 

After the model has run, there are many files that are present within the Model Folder that
can be viewed in-situ or loaded into R for viewing.

```{r gif-creation, gganimate = list(nframes = 30, fps = 3)}
# Rain file to grab the necessary rainfall method
rain_file <- desertHydro:::rainfallMethodCheck(ModelFolder, rainfall_method)
gifs <- desertHydro:::gifCreation(ModelFolder, rain_file = rain_file, discharge = F, saveGraph = T)
```

## Surface gif
```{r, gganimate = list(nframes = 30, fps = 3), fig.dim=c(6,8)}
gifs[[1]] # surface gif

```

## Velocity gif
```{r, gganimate = list(nframes = 30, fps = 3), fig.dim=c(6,8)}
gifs[[2]] # velocity gif
```




