---
title: "stream delineation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Purpose
The purpose of this article is to detail one of the most important steps in creating
a hydrological model: Watershed boundaries. A key component of watersheds is that
all of the water that falls within a watershed will end up at one point, commonly 
lakes, oceans, or within stream channels.

Commonly, watersheds are delineated or separated using digital elevation models or
DEMs. There are different algorithms to compute where water is going to go and that
is usually downhill or down in elevation. The most common algorithm is a D8 flow
direction that determines flow in one of 8 cardinal directions. After picking a
point or cell on a DEM, these algorithms will determine the cells that will 
eventually connect to that point.

# Creating a watershed boundary
There are many ways to create a watershed boundary with different GIS software.
You only need 3 things to create a watershed.

## 1) DEM
## 2) Pour point or outlet (point along the stream)
## 3) Flow algorithm (QGIS, Whitebox tools, ArcGIS Pro Spatial Analyst)

One important pre-processing step is to deal with any "pits" or "ridges" within
a DEM. A DEM is the average elevation over an area and if the spatial resolution
is coarse (not very detailed) then the flow path of water with get "stuck" within
the middle of the watershed instead of the outlet. 


```{r setup}
library(desertHydro)
```
